---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Volz Investigation

```{r}
source("MarkhovChain.R")
library("tidyverse")
```

## Simulating the virus

```{r}
num_sims = 1000
save_path = "SIR_virus_infections.RDS"

full_table <- markhov_virus(end_time = 11, beta = 0.7, gamma = 0.3, S=59, I=1, SIS=F)
full_table <- cbind(sim_num=1, full_table)

for (i in 2:num_sims){
  sim_table <- markhov_virus(end_time = 11, beta = 0.7, gamma = 0.3, S=59, I=1, SIS=F)
  sim_table <- cbind(sim_num=i, sim_table)
  full_table <- rbind(full_table, sim_table)
}

saveRDS(full_table, save_path)

```

## Do differential equations accurately estimate stochastic process? (SIR)

```{r}
source("Volz.R")
simulated_data <- readRDS("SIR_virus_infections.RDS")

get_mean_vals <- function(t, data){
  ids = unique(data$sim_num)
  Ss <- c()
  Is <- c()
  Rs <- c()
  for(i in 1:1000){
    run = subset(data, sim_num == i)
    most_recent_data_point = tail(subset(run, time <= t), n=1)
    # if (as.double(most_recent_data_point$I) > 0){
    Ss <- c(Ss, as.double(most_recent_data_point$S)) # Have to invoke as.double due to wierd formatting.
    Is <- c(Is, as.double(most_recent_data_point$I))
    Rs <- c(Rs, as.double(most_recent_data_point$R)) }
  # }
  return( c(S = mean(Ss), I = mean(Is), R = mean(Rs)) )
}
parameters <- c(beta = .7, gamma = .3)


SIRreal <- data.frame(time=c(), S=c(), I=c(), R=c())
for (t in seq(0,11, by=0.1)){
  mean_real_vals <- get_mean_vals(t, simulated_data)
  sim_row <- list(time=t, S=mean_real_vals[1]/60, I=mean_real_vals[2]/60, R=mean_real_vals[3]/60)
  SIRreal <- rbind(SIRreal, sim_row)
}
SIRreal

SIRoutput <- as.data.frame(ode(y = initial_state_sir, times = times, func = SIR_model, parms = parameters))

ggplot(data=SIRoutput, aes(x=time)) +
  # geom_smooth(aes(color="green", y=R)) +
  geom_smooth(aes(color="blue", y=I)) + 
  geom_smooth(aes(color="red", y=S)) +
  # geom_smooth(data=SIRreal, aes(color="darkgreen", x=time, y=R)) +
  geom_line(data=SIRreal, aes(color="lightblue", x=time, y=I),linewidth=1) +
  geom_line(data=SIRreal, aes(color="pink", x=time, y=S),linewidth=1) +
  scale_color_identity(name = "Function",
                       breaks = c("red", "pink", "blue", "lightblue"), #, "green", "darkgreen"),
                       labels = c("ODE S", "SIM S","ODE I", "SIM I"), #,"ODE R", "SIM R"),
                       aesthetics = c("color"),
                       guide="legend") +
  labs(title = "ODE v.s. Simulation")


```

## What about an SIS?

```{r}
source("Volz.R")
source("MarkhovChain.R")
simulated_data <- readRDS("SIS_virus_infections.RDS")
simulated_data
SISexpected <- data.frame(time=c(), I=c(), S=c())
for (t in seq(0,11, by=0.02)){
  expected_val <- sum(0:60*as.vector(markhov_probability(t,beta=0.7,gamma=0.3,initial_S = 59, initial_I = 1)))
  SISexpected <- rbind(SISexpected, list(time=t, I=expected_val/60, S=(60-expected_val)/60))
}

get_mean_vals <- function(t, data){
  ids = unique(data$sim_num)
  Ss <- c()
  Is <- c()
  for(i in ids){
    # print(j)
    # print(data)
    run = subset(data, sim_num == i)
    most_recent_data_point = tail(subset(run, time <= t), n=1)
    # if (as.double(most_recent_data_point$I) > 0){
    Ss <- c(Ss, as.double(most_recent_data_point$S)) # Have to invoke as.double due to wierd formatting.
    Is <- c(Is, as.double(most_recent_data_point$I))
  # }
    # print(Ss)
    # print(Is)
  }
  return( c(S = mean(Ss), I = mean(Is)))
} 
  
parameters <- c(beta = .7, gamma = .3)


SISreal <- data.frame(time=c(), S=c(), I=c())
for (t in seq(0,11, by=1)){
  mean_real_vals <- get_mean_vals(t, simulated_data)
  sim_row <- list(time=t, S=mean_real_vals[1]/60, I=mean_real_vals[2]/60)
  SISreal <- rbind(SISreal, sim_row)
}
SISreal

SISoutput <- as.data.frame(ode(y = initial_state_sis, times = times, func = SIS_model, parms = parameters))

ggplot(data=SISoutput, aes(x=time)) +
  geom_line(data=SISexpected, aes(color="blue4", x=time, y=I),linewidth=1) +
  geom_line(data=SISexpected, aes(color="red4", x=time, y=S),linewidth=1) +
  # geom_smooth(aes(color="green", y=R)) +
  geom_smooth(aes(color="blue", y=I)) + 
  geom_smooth(aes(color="red", y=S)) +
  # geom_smooth(data=SIRreal, aes(color="darkgreen", x=time, y=R)) +
  geom_line(data=SISreal, aes(color=alpha("lightblue",0.5), x=time, y=I), linewidth = 1) +
  geom_line(data=SISreal, aes(color=alpha("pink",0.5), x=time, y=S), linewidth = 1) +
  xlim(0,10) +
  scale_color_identity(name = "Function",
                       breaks = c("red", alpha("pink",0.5),"red4", "blue", alpha("lightblue",0.5),"blue4"), #, "green", "darkgreen"),
                       labels = c("ODE S", "SIM S","E[S]","ODE I", "SIM I", "E[I]"), #,"ODE R", "SIM R"),
                       aesthetics = c("color"),
                       guide="legend") +
  labs(title = "ODE v.s. Simulation v.s. Expected (SIS)")


```

## Plotting the likelihood function

```{r}
source("src/Volz.R")
simulated_data <- readRDS("src/SIS_virus_infections.RDS")
SIM_NUM = 200
run_data <- subset(simulated_data, sim_num == SIM_NUM)
n = nrow(run_data)
coalescent_times <- run_data[2:n,]
coalescent_times <- subset(coalescent_times, affected > 0)
coalescent_times <- subset(coalescent_times, time < 10)
coalescent_times <- coalescent_times$time
# coalescent_times

Volz_SIS(beta=0.7,gamma=0.3, sample_times = coalescent_times)

val_range <- seq(0, 1, by=0.05)
heatmap_vals <- data.frame(beta=c(), gamma=c(), likelihood=c())
for (i in val_range){
  for (j in val_range){
    heatmap_vals <- rbind(heatmap_vals, list(beta=i, gamma=j, likelihood=Volz_SIS(beta = i, gamma=j, sample_times = coalescent_times)))
  }
}

heatmap_vals_for_stats <- subset(heatmap_vals, likelihood != -Inf)
cutoff <- median(heatmap_vals_for_stats[,"likelihood"])
minim <- min(heatmap_vals_for_stats[,"likelihood"])
maxim <- max(heatmap_vals_for_stats[,"likelihood"])

SIS_like <- function(theta) -Volz_SIS(beta=theta[1],gamma=theta[2],sample_times = coalescent_times)
MLE <- optim(par = c(0.5, 0.5), fn= SIS_like, method="L-BFGS-B", lower = c(0.01,0.01), upper = c(1,1))
MLE$par

ggplot(heatmap_vals, aes(beta, gamma)) +
  geom_raster(aes(fill=likelihood)) +
  scale_fill_gradientn(colors=c("black", "midnightblue","white"),
                       breaks=c(minim,cutoff,maxim),
                       labels=c("min", "median","max")) +
  geom_point(aes(x=0.7, y=0.3),color="palevioletred3") +
    annotate('text', x=0.7, y=0.3 + 0.05, color="palevioletred3", label="True Value") +
    geom_point(aes(x=MLE$par[1], y=MLE$par[2]), color="seagreen") +
    annotate('text', x=MLE$par[1], y=MLE$par[2] + 0.05, label=expression("Estimate"), color="seagreen") +
    labs(
      title=paste("Volz Likelihood (2.3)"),
      x=expression("Beta"),
      y=expression("Gamma"),
      fill="Log Likelihood"
     
    )

```

### Volz Monte Carlo Instead of Diffirential Equation?

#### Compute the experimental A

```{r}
source("Volz.R")
simulated_data <- readRDS("SIS_virus_infections.RDS")
unique_sims <- unique(simulated_data$sim_num)
experimental_A <- data.frame()
for (i in unique_sims){
# for (i in 1:10){
  run_data <- subset(simulated_data, sim_num == i)
  # print(run_data)
  experimental_A <- rbind(experimental_A, A_from_data(times=times, sample_simulation = run_data, final_time = 10))
}

# experimental_A <- as.vector(experimental_A)
data <- colMeans(experimental_A,na.rm = T)
data <- as.vector(data)
saveRDS(data, file="experimental_A.RDS")
```

#### Plot experimental A vs ODE derived A

```{r}
source("Volz.R")
simulated_data <- readRDS("SIS_virus_infections.RDS")
experimental_A <- readRDS("experimental_A.RDS")
experimental_A
# experimental_A <- as.vector(experimental_A)
ODE_A <- SIS_A(seq(0,10,by=0.01))
ODE_A$A <- rev(ODE_A$A) # Account for working backwards in time
ODE_A

A_data <- data.frame(time=times, experimental=experimental_A)
ggplot(data=A_data) +
  xlim(0,10) +
  geom_line(color="red", aes(x=time,y=experimental),linewidth=1) +
  # geom_line(data=SISexpected,color="pink", aes(x=time,y=I),linewidth=1) +
  geom_line(data=ODE_A, color="blue",aes(x=time,y=A),linewidth=1) + 
  # geom_line(data=ODE_A, color="steelblue",aes(x=time,y=I),linewidth=1) +
  labs(y="fraction of population")
  
```

### Derivative of A
#### Bake Mixed A_derivative
```{r}
source("Volz.R")
simulated_data <- readRDS("SIS_virus_infections.RDS")
A_mixed <- A_dt_from_mix(times = seq(0,10,by=0.01), 
                         sample_simulation = simulated_data,
                         A_file = "experimental_A.RDS")
saveRDS(A_mixed, file="mixed_Adt.RDS")
```
#### Graph Mixed A v.s. experimental A
```{R}
source("Volz.R")
mixed_Adt <- readRDS("mixed_Adt.RDS")
data_Adt <- A_dt_from_data(A_file="experimental_A.RDS")
df <- data.frame(time = times, experimental=data_Adt, mixed=mixed_Adt)
ggplot(df, aes(x=time)) +
  geom_smooth(aes(y=experimental, color="#F8766D")) +
  geom_smooth(aes(y=mixed, color="#00BFC4")) +
  scale_color_identity(name = "Function",
                       breaks = c("#F8766D", "#00BFC4"),
                       labels = c("Approximated", "Derived"),
                       aesthetics = c("color"),
                       guide="legend") +
  labs(title = "Differential Mixed V/S Experimentally Approximated",
       )
```
#### Likelihood Function?
```{R}
data <- read.csv("simulation_results/monte_carlo_volz.csv")
data <- subset(data, simulation_used == 200)
maxs <- tail(data[order(replace_na(data$likelihood,-Inf)),],1)
heatmap_vals_for_stats <- subset(data, likelihood != -Inf)
cutoff <- median(heatmap_vals_for_stats[,"likelihood"])
minim <- min(heatmap_vals_for_stats[,"likelihood"])
maxim <- max(heatmap_vals_for_stats[,"likelihood"])
ggplot(data, aes(x=beta, y=gamma)) +
  geom_raster(aes(fill=likelihood)) +
   scale_fill_gradientn(colors=c("black", "midnightblue","white"),
                       breaks=c(minim,cutoff,maxim),
                       labels=c("min", "median","max")) +
  geom_point(aes(x=0.7, y=0.3),color="palevioletred3") +
    annotate('text', x=0.7, y=0.3 + 0.04, color="palevioletred3", label="True Value") +

  geom_point(aes(x=maxs$beta, y=maxs$gamma),color="seagreen") +
    annotate('text', x=maxs$beta, y=maxs$gamma + 0.04, color="seagreen", label="Estimate") +
    labs(
      title="Volz Monte Carlo Likelihood (2.4)",
      x="beta",
      y="gamma",
      fill="Log Likelihood"
    )

# A, Volz Sim study (higher n)
  
```
### Simulation Study (Volz, beta=0.7, gamma=0.3)
```{R}
ss_data <- read.csv(file="simulation_results/monte_carlo_volz.csv")
ss_data$X <- NULL
# ss_data
simulation_mle <- data.frame()
for (s in seq(1000)){
  dta <- subset(ss_data, simulation_used == s)
  dta <- dta[order(dta$likelihood,decreasing = T),]
  mle_row <- dta[1,]
  simulation_mle <- rbind(simulation_mle, mle_row)
}
mean(simulation_mle$beta)
mean(simulation_mle$gamma)

ggplot(data=simulation_mle, aes(fill=group)) +
  geom_histogram(aes(x=beta),fill="blue", position = "identity", alpha = 0.3, binwidth = 0.1) +
  geom_histogram(aes(x=gamma),fill="red", position = "identity", alpha = 0.3, binwidth = 0.1) + 
  scale_color_identity(name = "Parameter",
                       breaks = c("blue", "red"),
                       labels = c("Beta", "Gamma"),
                       aesthetics = c("fill"),
                       guide="legend") +
  # guides(fill = guide_legend(title = "Title")) +
  labs(
    title="Histograms for gamma and beta in sim. study",
    x="value"
  )
ggplot(data=simulation_mle, aes(x=beta, y=gamma)) +
  geom_point()



```